### 2014-09-23 paper/include/accessrecency.tex

\subsection{Smart Refresh\cite{smartrefresh} - from 2007}
The basic concept behind our scheme is that a memory row that has been recently read out or written to does not need to be refreshed again by the periodic refresh mechanism.

Uses a time-out counter for each memory row. The counter is set on a read/write/refresh and dicards refreshes until the timer hits zero. The counter is associated to a bank/row pair. The counters are stored in the memory controller. One counter consists of 2 SRAM bits. More bits give higher granularity and thus also better performance, the refresh is made less often.

This scheme will be most effective if all rows are accessed before a refresh is needed. If no accesses are made = the entire working set fits into cache, Smart Refresh becomes CBR (CAS before RAS). Only refresh one row at a time = inefficient, so they have made it possible to turn the technique off.

Implemented so only one row per each logical segments are deremented at the same time, which leads to that the refreshes are evenly distributed over time. (I.e. no burst refresh.)

@@

Our goal is to refresh only the data that will be used in the near future, and only refresh it if it is really needed. The other data is invalidated and/ or written back to main memory.

@@

For the cold lines four different (data-based) policies can be applied: all, valid, dirty or WB(n,m). "All" refreshes all lines. Valid refreshes the valid lines. Dirty refreshes the dirty lines.The WB policy is associated with a tuple (n, m). WB refreshes a Dirty line that is not being accessed for n times before writing it back and changing its state to Valid Clean; moreover, it refreshes a Valid Clean line that is not being accessed for m times before invalidating it. WB retains a Dirty line in the cache Ion ger because evicting it has the additional cost of writing the line back to lower-Ievel memory. To implement WB, we maintain a per-Iine Count. W hen the line is read or written, Count is set to n (if Dirty) or m (if Valid Clean). In addition, when the line is refreshed, Count is decremented. W hen Count reaches zero, the line is either written back or invalidated. Note that the Dirty policy is equivalent to WB(oo,O), while Valid is equivalent to WB(oo,oo). Finally, every policy refreshes cache Iines in transient states as well.

They have application categorization for the data-based policies (Cold ones). TODO

@@

Implementation: Cold lines. TODO

@@

Discuss energy savings.

>>>>

\subsection{Smart Refresh\cite{smartrefresh} - from 2007}
The basic concept behind our scheme is that a memory line that has been recently read out or written to does not need to be refreshed again by the periodic refresh mechanism.

Uses a time-out counter for each memory line. The counter is set on a read/write/refresh and dicards refreshes until the timer hits zero. The counter is associated to a bank/line pair. The counters are stored in the memory controller. One counter consists of 2 SRAM bits. More bits give higher granularity and thus also better performance because the refresh is made less often.

This scheme will be most effective if all lines are accessed before a refresh is needed. If no accesses are made = the entire working set fits into cache, Smart Refresh becomes CBR (CAS before RAS). Only refresh one line at a time = inefficient, so they have made it possible to turn the technique off.

Implemented so only one line per each logical segments are deremented at the same time, which leads to that the refreshes are evenly distributed over time. (I.e. no burst refresh.)

@@

Our goal is to refresh only the data that will be used in the near future, and only refresh if it is really needed. The other data is invalidated and/ or written back to main memory.

@@

For the cold lines four different (data-based) policies can be applied: \textit{all}, \textit{valid}, \textit{dirty} or WB(n,m). \textit{All} refreshes all lines. \textit{Valid} refreshes the valid lines. \textit{Dirty} refreshes the dirty lines. The WB policy is associated with a tuple (n, m). WB refreshes a Dirty line that is not being accessed for n times before writing it back and changing its state to Valid Clean; moreover, it refreshes a Valid Clean line that is not being accessed for m times before invalidating it. WB retains a Dirty line in the cache Ion ger because evicting it has the additional cost of writing the line back to lower-Ievel memory. To implement WB, we maintain a per-Iine Count. W hen the line is read or written, Count is set to n (if Dirty) or m (if Valid Clean). In addition, when the line is refreshed, Count is decremented. W hen Count reaches zero, the line is either written back or invalidated. Note that the Dirty policy is equivalent to WB(oo,O), while Valid is equivalent to WB(oo,oo). Finally, every policy refreshes cache lines in transient states as well.

They categorize applications based on how they are affected by the data-based policies. Fig. \ref{fig:app_cat} presents an application categorization based on two axes: application footprint and visibility. It is likely........ TODO

\begin{figure}[t!]
	\includegraphics[width=\textwidth/2]{app_cat}
	\caption{Application categorization according to the data-based policy \cite{refrint}.}
	\label{fig:app_cat}
\end{figure}

@@

Implementation: Cold lines. For each line there is a 5 bit counter and the line's state bits. Depending on the data policy  for each line, but they are somehow neglectible.  ...

Remember that this technique is targeted towards caches and eDRAM and only is tested in that enviroment. This means that we can not really use the results shown in their paper, but the techniques can be applied to DRAM.

@@

Discuss energy savings. Hard to do eDRAM vs DRAM...
